---
title: "Lesson 6: Graphing 2/2"
toc: true
fig-align: center
fig-width: 8
fig-height: 3.5
code-copy: true
code-overflow: wrap
code-line-numbers: true
code-fold: true
code-tools: true
eval: true
message: false
warning: false
# toc: true
---

```{r}
library(tidyverse)

dat <- read_csv("C:/Users/Bethany/Downloads/brighameaquinto.github.io-1/datasets/insurance.csv")
```


## Coloring 

- the color and fill arguments are accepted inside the aes() 


- color is the boarder color
- fill is the fill color

- you can set hard colors or set it to a column in your dataset
- R accepts color names in quotes and hex codes

- `scale_fill_manual()`
- `scale_color_manual()`



## Labeling 
#### `labs()`

**Explanation**

The `labs()` function is used to add label to the plot like the x-axis, the y-axis, and title. 

**Example**

```{r}
ggplot(dat, aes(x = region, fill = sex))+
    geom_bar()+ 
    labs(title = "Counts of Regions", x = "Regions", y = "Amount", fill = "Sex")+ theme_bw()
```


## Faceting
### `facet_wrap()` & `facet_grid()`
This of faceting as a third variable. We can set the x and y axes to see information, add color to make it

X Y: 
- color
- faceting

- `facet_wrap()` and `facet_grid()`


```{r}
#| fig.height = 5


color <- ggplot(dat, aes(x = region, fill = sex))+
            geom_bar()+ 
            labs(title = "Non-Faceted Graph", x = "", y = "Amount", fill = "Sex")+
            theme(
                title = element_text(size = 20), 
                axis.title = element_text(size = 10), 
                legend.title = element_text(size = 10))


facet <- ggplot(dat, aes(x = region, fill = sex))+
            geom_bar()+ 
            facet_wrap(~sex)+
            labs(title = "Faceted Graph", x = "", y = "Amount", fill = "Sex")+
            theme(
                title = element_text(size = 20), 
                axis.title = element_text(size = 10), 
                legend.title = element_text(size = 10), 
                legend.position = "none")




ggpubr::ggarrange(color, facet, nrow = 2)
```



## Themes 

The following theme options affect all the non-data aspects

- `theme_bw()`
- `theme_linedraw()`
- `theme_light()`
- `theme_dark()`
- `theme_minimal()`
- `theme_classic()`
- `theme_void()`
- `theme_test()`





Next week we'll get into colors, gg_text()/gg_label(), labs(), themes(), legend?, facet_wrap(), 
geom_boxplot() is often used with geom_jitter() to make powerful visualizations.