---
title: "Lesson 4: Last Three Wrangling Verbs"
toc: true
---


Before we do any exercises, make sure to run this code: 

```{r}
#| message: false
#| warning: false
#| code-overflow: wrap

library(tidyverse)
dat <- read_csv("https://raw.githubusercontent.com/BrighamEaquinto/brighameaquinto.github.io/master/datasets/insurance.csv")
```

This is our dataset for this practice: 

```{r}
#| code-fold: true
#| include: false

# head(dat) %>% knitr::kable()
```

> | age|sex    |    bmi| children|smoker |region    |   charges|
> |---:|:------|------:|--------:|:------|:---------|---------:|
> |  19|female | 27.900|        0|yes    |southwest | 16884.924|
> |  18|male   | 33.770|        1|no     |southeast |  1725.552|
> |  28|male   | 33.000|        3|no     |southeast |  4449.462|
> |  33|male   | 22.705|        0|no     |northwest | 21984.471|
> |  32|male   | 28.880|        0|no     |northwest |  3866.855|

<br>

## Let's learn about `group_by()` & `summarise()` 


What if we want to learn: 

- the average BMI ***by*** gender 
- the total sum of charges ***by*** region in the US 
- min & max of children ***by *** age

All of these are examples uses for `group_by()` and `summarise()` statements. 

<br>

# Examples 

## Examples 1: Average BMI by Gender

```{r}
#| eval: false

dat %>% 
    group_by(sex) %>%
    summarise(Average_Age = mean(age)) 
```

> |sex    | Average_Age|
> |:------|-----------:|
> |female |    39.50302|
> |male   |    38.91716|

## Examples 2: Average BMI by Age

```{r}
#| eval: false 

dat %>% 
    group_by(age) %>% 
    summarise(mean(bmi)) %>% 
    head() 
```

> | age| mean(bmi)|
> |---:|---------:|
> |  18|  31.32616|
> |  19|  28.59691|
> |  20|  30.63276|
> |  21|  28.18571|
> |  22|  31.08768|
> |  23|  31.45446|


# Practice Exercises

## Exercise 1: Sum of Charges by Region

Use the code below to get you started. Find the sum of the charges by region. 

**Starter Code:**

```{r}
#| eval: false
dat %>% 
    group_by(______) %>% 
    summarise(_____________ = ___(_______))
```


```{r}
#| echo: false
#| include: false
dat %>% 
    group_by(region) %>% 
    summarise(Total_Charges = sum(charges)) %>% 
    knitr::kable()
```


Attempt the code above then use this dropdown for help

```{r}
#| code-fold: true
#| code-summary: "Answer"
#| eval: false
dat %>% 
    group_by(region) %>% 
    summarise(Total_Charges = sum(charges))
```

This is the correct output: 

> |region    | Total_Charges|
> |:---------|-------------:|
> |northeast |       4343669|
> |northwest |       4035712|
> |southeast |       5363690|
> |southwest |       4012755|


## Exercise 2: Min & Max of Children by Age

Use the code below to get you started. Find the minimum and maximum amount of children by age. 

**Starter Code:**

```{r}
#| eval: false
dat %>% 
    group_by(___) %>%
    summarise(
        Min = ___(________),
        Max = ___(________)
        ) %>% 
    head()

```

```{r}
#| echo: false
#| include: false
dat %>% 
    group_by(age) %>%
    summarise(
        Min = min(children),
        Max = max(children)
        ) %>% 
    head() %>% 
    knitr::kable()

```

```{r}
#| code-fold: true
#| code-summary: "Answer"
#| eval: false

dat %>% 
    group_by(age) %>%
    summarise(
        Min = min(children),
        Max = max(children)
        ) %>% 
    head() 

```

This is the correct output: 

> | age| Min| Max|
> |---:|---:|---:|
> |  18|   0|   4|
> |  19|   0|   5|
> |  20|   0|   5|
> |  21|   0|   4|



## Exercise 3: Average Charges by Sex and by Smoker

Here no starter code is given. If you need to, look back at previous code chunks for guidence. 

Try as best as possible to mimic the results below. 

Only use the dropdown answer when you've tried as much as possible.

```{r}
#| code-fold: true
#| code-summary: "Answer"
#| eval: false

dat %>% 
    group_by(sex, smoker) %>% 
    summarise(
        Average_Charges = mean(charges), 
        Total_Charges = sum(charges)) 
```

This is the correct output: 

> |sex    |smoker | Average_Charges| Total_Charges|
> |:------|:------|---------------:|-------------:|
> |female |no     |        8762.297|       4792977|
> |female |yes    |       30678.996|       3528085|
> |male   |no     |        8087.205|       4181085|
> |male   |yes    |       33042.006|       5253679|


## Exercise 4: Storm Dataset - Wind and Air Pressure

This dataset is bult into R so you already have it. It is called `storms`. This is a new dataset so you may want to either `view()` it or use the `head()` method on it.

Find the average and standard deviation of the wind speed and air pressure by storm category.

 
```{r}
#| code-fold: true
#| code-summary: "Answer"
#| eval: false

storms %>% 
    group_by(category) %>% 
    summarise(
        mean(wind),
        sd(wind), 
        mean(pressure), 
        sd(pressure)
        ) %>% knitr::kable()

```


This is the correct output: 

> |category | mean(wind)| sd(wind)| mean(pressure)| sd(pressure)|
> |:--------|----------:|--------:|--------------:|------------:|
> |-1       |   27.49482| 3.702448|      1007.5390|     3.900209|
> |0        |   45.66392| 8.231625|       999.2910|     6.837831|
> |1        |   70.95140| 5.454009|       981.1887|     9.125784|
> |2        |   89.41923| 3.873391|       966.9359|     9.018024|
> |3        |  104.48157| 4.246545|       953.9124|     9.168259|
> |4        |  122.10462| 6.383000|       939.3942|     9.788015|
> |5        |  145.58140| 5.862072|       917.4070|    12.732892|



## Exercise 5: Summarizing by Month

Find the mean, median, and standard deviation of wind speed by month. 

Only use the dropdown answer when you've tried as much as possible.

```{r}
#| code-fold: true
#| code-summary: "Answer"
#| eval: false

storms %>% 
    group_by(month) %>% 
    summarise(
        Mean_Wind = mean(wind), 
        Median = median(wind), 
        St_Dev = sd(wind)
        ) 
```

This is the correct output: 

> | month| Mean_Wind| Median|    St_Dev|
> |-----:|---------:|------:|---------:|
> |     1|  50.33333|     50| 12.312969|
> |     4|  44.61538|     45|  5.937711|
> |     5|  36.27778|     35|  9.011887|
> |     6|  37.55348|     35| 12.676257|
> |     7|  41.74973|     40| 18.711162|
> |     8|  51.75471|     45| 25.584917|
> |     9|  57.54597|     50| 28.192273|
> |    10|  56.49272|     50| 25.795906|
> |    11|  53.30986|     50| 22.702523|
> |    12|  47.88000|     45| 14.608659|



## Exercise 6: Airquality Dataset - Summarising by Month

This example is very similar to the previous example. This is a new datset and hopefully doing the same thing with another dataset will solidify the concept. 

This dataset is bult into R so you already have it. It is called `airquality`. This is a new dataset so you may want to either `view()` it or use the `head()` method on it.


```{r}
#| code-fold: true
#| code-summary: "Answer"
#| eval: false

airquality %>% 
    group_by(Month) %>% 
    summarise(
        Mean_Wind = mean(Wind), 
        Median = median(Wind), 
        St_Dev = sd(Wind)
        ) 
```

This is the correct output: 

> | Month| Mean_Wind| Median|   St_Dev|
> |-----:|---------:|------:|--------:|
> |     5| 11.622581|   11.5| 3.531450|
> |     6| 10.266667|    9.7| 3.769234|
> |     7|  8.941935|    8.6| 3.035981|
> |     8|  8.793548|    8.6| 3.225930|
> |     9| 10.180000|   10.3| 3.461254|