---
title: "Lesson 2: Reading Data"
subtitle: Learning the ggplot2 package
toc: true
---

## Looking At Datasets

Options to See Data:

- `view()`
- `head()`

<br>

## Build-in Datasets

Run this code to see a list of built in datasets: 
```{r}
#| eval: false
data()
```

The `iris` dataset is commonly used and built into R; no additional downloads are necessary. 

```{r}
#| eval: false
head(iris)
```
```{r}
#| eval: false
view(iris)
```
<br>

## Reading Data from URL

Read in this dataset from this URL: https://raw.githubusercontent.com/saundersg/Statistics-Notebook/master/Data/Rent.csv 

```{r}
#| eval: false
housing_data <- read_csv("https://raw.githubusercontent.com/saundersg/Statistics-Notebook/master/Data/Rent.csv")

head(housing_data)

```
<br> 

## Reading Data from Local CSV File

First, let's make a folder on your computer to use for bootcamp. I suggest putting it in your downloads folder or your documents folder and call it "R Bootcamp". 

Next, click the "Session" button at the top of RStudio, then "Set Working Directory", then "Choose Directory" and use your computer's file directory to navigate to the folder you made for bootcamp. 

This set your "working directory" which basically makes it easier for us to deal with files. 



<p>Download the <a href="datasets\insurance.csv" target="_blank">datasets\insurance.csv</a> document to provide your answers to questions asked in the steps below.</p>


# Other Topics: 

- directories
- running code