---
title: "Week 4: Grouping and Summarising"
date: October 19
toc: true
toc-depth: 5
toc-title: "Outline"
---


Before we do any exercises, make sure to run this code: 

```{r}
#| message: false
#| warning: false
#| code-overflow: wrap

library(tidyverse)
dat <- read_csv("https://raw.githubusercontent.com/BrighamEaquinto/brighameaquinto.github.io/master/datasets/insurance.csv")
```

This is our dataset for this practice: 

```{r}
#| code-fold: true
#| include: false

# head(dat) %>% knitr::kable()
```

| age|sex    |    bmi| children|smoker |region    |   charges|
|---:|:------|------:|--------:|:------|:---------|---------:|
|  19|female | 27.900|        0|yes    |southwest | 16884.924|
|  18|male   | 33.770|        1|no     |southeast |  1725.552|
|  28|male   | 33.000|        3|no     |southeast |  4449.462|
|  33|male   | 22.705|        0|no     |northwest | 21984.471|
|  32|male   | 28.880|        0|no     |northwest |  3866.855|

<br>

## Let's learn about `group_by() %>% summarise()` 


What if we want to learn: 

- the average BMI ***by*** gender 
- the total sum of charges ***by*** region in the US 
- min & max of children ***by *** age

All of these are examples uses for `group_by()` and `summarise()` statements. 

<br>

# Examples 

## Exercise 1: Average BMI by Gender

```{r}
dat %>% 
    group_by(sex) %>%
    summarise(Average_Age = mean(age))
```

## Exercise 2: Average BMI by Age

```{r}
dat %>% 
    group_by(age) %>% 
    summarise(mean(bmi)) %>% 
    head() 
```


# Your Turn

## Exercise 3: Sum of Charges by Region

Use the code below to get you started. Find the sum of the charges by region. 

**Starter Code:**

```{r}
#| eval: false
dat %>% 
    group_by() %>% 
    summarise()
```

This is the correct output
```{r}
#| echo: false
dat %>% 
    group_by(region) %>% 
    summarise(Total_Charges = sum(charges))
```

Attempt the code above then use this dropdown for help

```{r}
#| code-fold: true
#| code-summary: "Answer"
#| eval: false
dat %>% 
    group_by(region) %>% 
    summarise(Total_Charges = sum(charges))
```


## Exercise 4: Min & Max of Children by Age

```{r}
dat %>% 
    group_by(age) %>%
    summarise(
        Min = min(children),
        Max = max(children)
        ) %>% 
    head()

```


## Exercise 5: Average Charges by Sex and by Smoker

```{r}
dat %>% 
    group_by(sex, smoker) %>% 
    summarise(
        Average_Charges = mean(charges), 
        Total_Charges = sum(charges))
```


